---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# API and Github

Here is a link to the github repo[https://github.com/jzemmels/STAT585_lab4_group10]

Next we have inlcuded code that could take data from the API thanks to Stephanie!

```{r echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(rvest)
library(jsonlite)
library(maps)
library(shiny)
```


```{r api-data}
# Read the API Data
url <- "https://data.iowa.gov/resource/m3tr-qhgy.json?county=Story"
d <- jsonlite::fromJSON(url)
```

```{r naming-convention}
# Change column names to match Iowa_Liquor_Sales-Story.csv

names(d)[names(d) == 'invoice_line_no'] <- 'Invoice.Item.Number'
names(d)[names(d) == 'name'] <- 'Store.Name'
names(d)[names(d) == 'zipcode'] <- 'Zip.Code'
names(d)[names(d) == 'county'] <- 'County'
names(d)[names(d) == 'vendor_no'] <- 'Vendor.Number'
names(d)[names(d) == 'vendor_name'] <- 'Vendor.Name'
names(d)[names(d) == 'state_bottle_cost'] <- 'State.Bottle.Cost'
names(d)[names(d) == 'sale_dollars'] <- 'Sale..Dolars.'
names(d)[names(d) == 'date'] <- 'Date'
names(d)[names(d) == 'address'] <- 'Address'
names(d)[names(d) == 'category'] <- 'Category'
names(d)[names(d) == 'im_desc'] <- 'Item.Description'
names(d)[names(d) == 'pack'] <- 'Pack'
names(d)[names(d) == 'state_bottle_retail'] <- 'State.Bottle.Retail'
names(d)[names(d) == 'state_bottle_retail'] <- 'Volume.Sold..Liters.'
names(d)[names(d) == 'store'] <- 'Store.Number'
names(d)[names(d) == 'store_location_city'] <- 'City'
names(d)[names(d) == 'county_number'] <- 'County.Number'
names(d)[names(d) == 'category_name'] <- 'Category.Name'
names(d)[names(d) == 'itemno'] <- 'Item.Number'
names(d)[names(d) == 'bottle_volume_ml'] <- 'Bottle.Volume..ml.'
names(d)[names(d) == 'store'] <- 'Store.Number'
names(d)[names(d) == 'sale_gallons'] <- 'Volume.Sold..Gallons.'
names(d)[names(d) == 'sale_bottles'] <- 'Bottles.Sold'
names(d)[names(d) == 'sale_liters'] <- 'Volume.Sold..Liters.'

```

```{r format-api}
# Create and format Store.Location variable to match Iowa_Liquor_Sales-Story.csv
dflat <- flatten(flatten(d))
store_location_lat <- as.numeric(unlist(as.character(purrr::map(dflat$store_location.coordinates, function(d) d[[1]]))))

store_location_long <- as.numeric(unlist(as.character(purrr::map(dflat$store_location.coordinates, function(d) d[[2]]))))

d <- cbind(d, store_location_lat, store_location_long)

d <- d %>% mutate(city.zip = paste(city,Zip.Code),
             lat.lon = paste0("(",store_location_lat,', ',store_location_long,")"),
             Store.Location = factor(paste(Address,city.zip,lat.lon,sep="\n")))
```

```{r api-reorder}

# Reorder columns to match Iowa_Liquor_Sales-Story.csv
d <- d %>% select(Invoice.Item.Number,
             Date,
             Store.Number,
             Store.Name,
             Address,
             City,
             Zip.Code,
             Store.Location,
             County.Number,
             County,
             Category,
             Category.Name,
             Vendor.Number,
             Vendor.Name,
             Item.Number,
             Item.Description,
             Pack,
             Bottle.Volume..ml.,
             State.Bottle.Cost,
             State.Bottle.Retail,
             Bottles.Sold,
             Sale..Dolars.,
             Volume.Sold..Liters.,
             Volume.Sold..Gallons.)

```


```{r api-change-class}

# Change variable classes to match Iowa_Liquor_Sales-Story.csv
d <- d %>% mutate(Invoice.Item.Number = factor(Invoice.Item.Number),
             Date = factor(Date),
             Store.Number = as.integer(Store.Number),
             Store.Name = factor(Store.Name),
             Address = factor(Address),
             City = factor(City),
             Zip.Code = as.integer(Zip.Code),
             County.Number = as.integer(County.Number),
             County = factor(County),
             Category = as.integer(Category),
             Category.Name = factor(Category.Name),
             Vendor.Number = as.integer(Vendor.Number),
             Item.Number = as.integer(Item.Number),
             Pack = as.integer(Pack),
             Bottle.Volume..ml. = as.integer(Bottle.Volume..ml.),
             State.Bottle.Cost = as.double(State.Bottle.Cost),
             State.Bottle.Retail = as.double(State.Bottle.Retail),
             Bottles.Sold = as.integer(Bottles.Sold),
             Sale..Dolars. = as.double(Sale..Dolars.),
             Volume.Sold..Liters. = as.double(Volume.Sold..Liters.),
             Volume.Sold..Gallons. = as.double(Volume.Sold..Gallons.))
```

```{r save-api}
# Save data to data folder
# write.csv(d,'data/Iowa_Liquor_Sales-Story_1000.csv')
```

# Spatial Cleaning

First we read in our data provided from the lab assignment. This portion of code
was created by Joe and Charlotte. 

```{r read-dat}
# Read in prepared data
liquor <- read_csv("data/story-sales/Iowa_Liquor_Sales-Story.csv")
```

Separate the lat/long data and make a new column for type of liquor (Charlotte). 
There were many different names for different types of alcohol so we simplified
this category down to five different entries. 

```{r lat-lon}
# Extract location data and find type of alcohol 
liquor <- liquor %>%
  mutate(coord.data = str_extract(`Store Location`, "\\(.*\\)"),
         lat = str_extract(coord.data, "\\(.*,"),
         lon = str_extract(coord.data, ",.*\\)")) %>%
  mutate(lat = str_extract(lat, "[0-9\\.-]{1,}"),
         lon = str_extract(lon, "[0-9\\.-]{1,}")) %>%
  mutate(lat = as.numeric(lat),
         lon = as.numeric(lon),
         `Category Name` = str_to_lower(`Category Name`, locale = "en"),
         category.type = case_when(
           str_detect(`Category Name`, "whisk") ~ "whiskey",
           str_detect(`Category Name`, "vodka") ~ "vodka",
           str_detect(`Category Name`, "rum") ~ "rum",
           str_detect(`Category Name`, "brand") ~ "brandy",
           TRUE ~ "other"
         ))
```

We also would like to visualize sales by brand of liquor (Joe). There were too many individual brands, so we settled on a handful of recognizable names and grouped the rest.

```{r liquor-brand}
liquor <- liquor %>%
  mutate(brand = case_when(
    str_detect(`Item Description`,"[Bb]acardi") ~ "Bacardi",
    str_detect(`Item Description`,"[Aa]bsolut") ~ "Absolut",
    str_detect(`Item Description`,"[Ss]mirnoff") ~ "Smirnoff",
    str_detect(`Item Description`,"[Ss]vedka") ~ "Svedka",
    str_detect(`Item Description`,"[Jj]ameson") ~ "Jameson",
    str_detect(`Item Description`,"[Cc]aptain [Mm]organ") ~ "Captain Morgan",
    str_detect(`Item Description`,"[Tt]anqueray") ~ "Tanqueray",
    str_detect(`Item Description`,"[Rr]umchata") ~ "Rumchata",
    str_detect(`Item Description`,"[Hh]ennessy") ~ "Hennessy",
    str_detect(`Item Description`,"U[Vv]") ~ "UV",
    str_detect(`Item Description`,"[Jj]ameson") ~ "Jameson",
    str_detect(`Item Description`,"[Jj]ack [Dd]aniels") ~ "Jack Daniels",
    str_detect(`Item Description`,"[Mm]alibu") ~ "Malibu",
    str_detect(`Item Description`,"[Ss]outhern [Cc]omfort") ~ "Southern Comfort",
    str_detect(`Item Description`,"[Tt]itos") ~ "Titos",
    str_detect(`Item Description`,"[Ff]ireball") ~ "Fireball",
    str_detect(`Item Description`,"[Jj]agermeister") ~ "Jagermeister",
    str_detect(`Item Description`,"[Pp]aramount") ~ "Paramount",
    str_detect(`Item Description`,"[Pp]hillips") ~ "Phillips",
    str_detect(`Item Description`,"[Ss]eagrams") ~ "Seagrams",
    TRUE ~ "Other"
  ))
```

County had both "Story" and "STORY" so we standardize below

```{r standard-county}
# Standardize county name
liquor <- liquor %>%
  mutate(county = "Story")
```

Story county map so that we can plot the lat/long of all of Story county.

```{r make-map}
county <- map("county","iowa.story")
```


We decided to drop a lot of the columns to reduce the dataset. We may choose to pick some back up at a later date, but this should help with the size of the data.

```{r clean-spatial}
# Remove extraneous columns
liquor_clean <- liquor %>%
  select(-c(`Invoice/Item Number`,`Store Number`,Address,`Zip Code`,`Store Location`,`County Number`,County,`Category Name`,`Vendor Number`,`Item Number`,coord.data,Category,`Item Description`,`Volume Sold (Gallons)`,`Volume Sold (Liters)`,`Sale (Dollars)`,`Bottles Sold`,`State Bottle Retail`,`State Bottle Cost`,`Bottle Volume (ml)`,Pack))

# write_csv(liquor_clean,"data/liquor_cleanForSpatial.csv",col_names = TRUE)
```


# Temporal Cleaning

This portion of code was created by Yuchen and Hao. In order to analysis the dataset in temporal aspect, we transformed `Date` variable into week (Monday, Tuesday, ...), month, quarter, and year four new variables.

```{r change-date}
# Transform date to character string

liquor$Date=as.character(liquor$Date)

# Summary

liquor$Date %>% mdy() %>% year() %>% table()

# Create new temporal variables

liquor$Date=liquor$Date %>% mdy
liquor$week=liquor$Date %>% wday() %>% factor()
levels(liquor$week)= c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")
liquor$month=liquor$Date %>% month()
liquor$quarter=liquor$Date %>% quarter()
liquor$year=liquor$Date %>% year()

```

Check interested response variable

```{r summarise-vars}
liquor$Bottles.Sold %>% summary
liquor$Sale..Dollars. %>% summary
```

We decided to drop a lot of the columns to reduce the dataset. We may choose to pick some back up at a later date, but this should help with the size of the data.

```{r clena-temporal}
# Create new data frame with only relevant variables
liquor_clean <- liquor %>%
  select(c(Date, `Bottles Sold`, `Sale (Dollars)`,brand,week,month,quarter,year))
class(liquor_clean)

# Write to csv
# write_csv(liquor_clean,"data/liquor_cleanForTemporal.csv",col_names = TRUE)
```

Please see our github to run the app...
