---
title: "R Notebook"
output: html_notebook
---

Read raw data directly from .zip file 

```{r}
library(tidyverse)

liquor <- read.csv(unz("Data/story-sales.zip", "Iowa_Liquor_Sales-Story.csv"))
```

We (Yuchen & Hao) utilized Joe and charlotte's code to create new variable brand
```{r}
liquor <- liquor %>%
  mutate(brand = case_when(
    str_detect(`Item.Description`,"[Bb]acardi") ~ "Bacardi",
    str_detect(`Item.Description`,"[Aa]bsolut") ~ "Absolut",
    str_detect(`Item.Description`,"[Ss]mirnoff") ~ "Smirnoff",
    str_detect(`Item.Description`,"[Ss]vedka") ~ "Svedka",
    str_detect(`Item.Description`,"[Jj]ameson") ~ "Jameson",
    str_detect(`Item.Description`,"[Cc]aptain [Mm]organ") ~ "Captain Morgan",
    str_detect(`Item.Description`,"[Tt]anqueray") ~ "Tanqueray",
    str_detect(`Item.Description`,"[Rr]umchata") ~ "Rumchata",
    str_detect(`Item.Description`,"[Hh]ennessy") ~ "Hennessy",
    str_detect(`Item.Description`,"U[Vv]") ~ "UV",
    str_detect(`Item.Description`,"[Jj]ameson") ~ "Jameson",
    str_detect(`Item.Description`,"[Jj]ack [Dd]aniels") ~ "Jack Daniels",
    str_detect(`Item.Description`,"[Mm]alibu") ~ "Malibu",
    str_detect(`Item.Description`,"[Ss]outhern [Cc]omfort") ~ "Southern Comfort",
    str_detect(`Item.Description`,"[Tt]itos") ~ "Titos",
    str_detect(`Item.Description`,"[Ff]ireball") ~ "Fireball",
    str_detect(`Item.Description`,"[Jj]agermeister") ~ "Jagermeister",
    str_detect(`Item.Description`,"[Pp]aramount") ~ "Paramount",
    str_detect(`Item.Description`,"[Pp]hillips") ~ "Phillips",
    str_detect(`Item.Description`,"[Ss]eagrams") ~ "Seagrams",
    TRUE ~ "Other"
  ))

```

In order to analysis the dataset in temporal aspect, we transformed `Date` variable into week (Monday, Tuesday, ...), month, quarter, and year four new variables.
```{r}
library(lubridate)
liquor$Date=as.character(liquor$Date)

# summary
liquor$Date %>% mdy() %>% year() %>% table()

# create new temporal variables
liquor$week=liquor$Date %>% mdy() %>% wday() %>% factor()
levels(liquor$week)= c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")
liquor$month=liquor$Date %>% mdy() %>% month()
liquor$quarter=liquor$Date %>% mdy() %>% quarter()
liquor$year=liquor$Date %>% mdy() %>% year()

```

Check interested response variable

```{r}
liquor$Bottles.Sold %>% summary
liquor$Sale..Dollars. %>% summary
```

We decided to drop a lot of the columns to reduce the dataset. We may choose to pick some back up at a later date, but this should help with the size of the data.

```{r}
liquor_clean <- liquor %>%
  select(-c(`Invoice.Item.Number`,`Store.Number`,Address,`Zip.Code`,`Store.Location`,`County.Number`,County,`Vendor.Number`,`Item.Number`))
class(liquor_clean)
write_csv(liquor_clean,"data/liquor_cleanForTemporal.csv",col_names = TRUE)
```

