---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(maps)

liquor <- read_csv("data/story-sales/Iowa_Liquor_Sales-Story.csv")
```

Separate the lat/long data and make a new column for type of liquor (Charlotte):

```{r}
liquor <- liquor %>%
  mutate(coord.data = str_extract(`Store Location`, "\\(.*\\)"),
         lat = str_extract(coord.data, "\\(.*,"),
         lon = str_extract(coord.data, ",.*\\)")) %>%
  mutate(lat = str_extract(lat, "[0-9\\.-]{1,}"),
         lon = str_extract(lon, "[0-9\\.-]{1,}")) %>%
  mutate(lat = as.numeric(lat),
         lon = as.numeric(lon),
         `Category Name` = str_to_lower(`Category Name`, locale = "en"),
         category.type = case_when(
           str_detect(`Category Name`, "whisk") ~ "whiskey",
           str_detect(`Category Name`, "vodka") ~ "vodka",
           str_detect(`Category Name`, "rum") ~ "rum",
           str_detect(`Category Name`, "brand") ~ "brandy",
           TRUE ~ "other"
         ))

# liquor %>%
#   group_by(category.type) %>%
#   summarise(count=n())
```

We also would like to visualize sales by brand of liquor (Joe). There were too many individual brands, so we settled on a handful of recognizable names and grouped the rest.

```{r}
liquor <- liquor %>%
  mutate(brand = case_when(
    str_detect(`Item Description`,"[Bb]acardi") ~ "Bacardi",
    str_detect(`Item Description`,"[Aa]bsolut") ~ "Absolut",
    str_detect(`Item Description`,"[Ss]mirnoff") ~ "Smirnoff",
    str_detect(`Item Description`,"[Ss]vedka") ~ "Svedka",
    str_detect(`Item Description`,"[Jj]ameson") ~ "Jameson",
    str_detect(`Item Description`,"[Cc]aptain [Mm]organ") ~ "Captain Morgan",
    str_detect(`Item Description`,"[Tt]anqueray") ~ "Tanqueray",
    str_detect(`Item Description`,"[Rr]umchata") ~ "Rumchata",
    str_detect(`Item Description`,"[Hh]ennessy") ~ "Hennessy",
    str_detect(`Item Description`,"U[Vv]") ~ "UV",
    str_detect(`Item Description`,"[Jj]ameson") ~ "Jameson",
    str_detect(`Item Description`,"[Jj]ack [Dd]aniels") ~ "Jack Daniels",
    str_detect(`Item Description`,"[Mm]alibu") ~ "Malibu",
    str_detect(`Item Description`,"[Ss]outhern [Cc]omfort") ~ "Southern Comfort",
    str_detect(`Item Description`,"[Tt]itos") ~ "Titos",
    str_detect(`Item Description`,"[Ff]ireball") ~ "Fireball",
    str_detect(`Item Description`,"[Jj]agermeister") ~ "Jagermeister",
    str_detect(`Item Description`,"[Pp]aramount") ~ "Paramount",
    str_detect(`Item Description`,"[Pp]hillips") ~ "Phillips",
    str_detect(`Item Description`,"[Ss]eagrams") ~ "Seagrams",
    TRUE ~ "Other"
  ))

# liquor %>%
#   group_by(brand) %>%
#   summarise(count=n())
```

County had both "Story" and "STORY"

```{r}
liquor <- liquor %>%
  mutate(county = "Story")
```

Story county map so that we can plot the lat/long of all of Story county.
```{r}
county <- map("county","iowa.story")
```

```{r}
countyMap <- data.frame(lat=county$y,long=county$x) %>% na.omit()
countyMap <- data.frame(lat=county$x,long=county$y) %>% na.omit()
write_csv(countyMap,path = "data/storyCountyMap.csv",col_names = TRUE)
countyMap %>%
  na.omit() %>%
  ggplot(aes(x=long,y=lat)) +
  geom_path()

liquor %>%
  ggplot(aes(x = lon, y = lat))+
  geom_point()
```

We decided to drop a lot of the columns to reduce the dataset. We may choose to pick some back up at a later date, but this should help with the size of the data.

```{r}
liquor_clean <- liquor %>%
  select(-c(`Invoice/Item Number`,`Store Number`,Address,`Zip Code`,`Store Location`,`County Number`,County,`Category Name`,`Vendor Number`,`Item Number`,coord.data,Category,`Item Description`,`Volume Sold (Gallons)`,`Volume Sold (Liters)`,`Sale (Dollars)`,`Bottles Sold`,`State Bottle Retail`,`State Bottle Cost`,`Bottle Volume (ml)`,Pack))

# write_csv(liquor_clean,"data/liquor_cleanForSpatial.csv",col_names = TRUE)
```

